---
title: "R example calls"
output: pdf_document
date: "2025-01-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set working directory and load in api methods
```{r}
# library(roxygen2)
# roxygen2::load_pkgload("./FEHreadR")
# OR
# install.packages("devtools")
# devtools::install_github("UI-Research/RreadFEH/R/FEHreadR")
library(FEHreadR)
```

Connect to server
```{r}
email <- "email@urban.org"
password <- "password123"

con <- dm_connect(email, password)
```

Get projects
```{r}
projects <- dm_get_projects(con)
print(projects)
```

Get a specific project
```{r}
project_id <- 1
project_info <- dm_get_project(con, project_id)
print(project_info)
```

Get scenarios
```{r}
scenarios <- dm_get_scenarios(con)
print(scenarios)
```

Get scenarios for a project by project id
```{r}
project_id <- 1
scenarios1 <- dm_get_scenarios_for_project(con, project_id)
print(scenarios1)
```

Get variables for a project by project id
```{r}
project_id <- 1
variables <- dm_get_variables_for_project(con, project_id)

family_variables <- as.data.frame(variables[1, 'var_definitions'])
person_variables <- as.data.frame(variables[2, 'var_definitions'])

print(family_variables)
print(person_variables)
```

Request a subset of data
```{r}
project_name = "BabyBonds"
scenario_names = list("Baseline_v1", "BabyBonds_v1")
person_variables = list("perid", "race", "sex", "year_died", "earnings")
family_variables = list("fam_id", "numkids", "immigration_year")
birth_year_range = list(1950, 2100) # optional
year_range = list(1950, 2100) # optional

response <- dm_generate_dataset(
  con,
  project_name=project_name,
  scenarios=scenario_names,
  family_variables=family_variables,
  person_variables=person_variables,
  birth_year_range=birth_year_range,
  year_range=year_range
)

print(response)

# Retrieve job_id from response
job_id <- response$job_id
```

Get status of the requested dataset
```{r}
file_type <- "csv" # only `csv` or `parquet` allowed

response <- dm_get_dataset_status(con, job_id, file_type)

# Get family and person presigned urls once job_status=="SUCCEEDED"
family_url <- response$family_url
person_url <- response$person_url
```


Download files locally
```{r}
dm_download_file(family_url, "/path/to/local/downloads/family_file.zip") # only .zip or .pq
dm_download_file(person_url, "/path/to/local/downloads/person_file.zip") # only .zip or .pq
```


Request and download datasets
Calls the above functions and saves the data as family_data.{ext} or person_data.{ext}
```{r}
project_name = "BabyBonds"
scenario_names = list("Baseline_v1", "BabyBonds_v1")
person_variables = list("perid", "race", "sex", "year_died", "earnings")
family_variables = list("fam_id", "numkids", "immigration_year")
birth_year_range = list() # optional
year_range = list() # optional

file_type = "csv" # `csv` or `parquet` only
output_dir = "/users/local/downloads"

dm_request_and_download_datasets(
    con,
    file_type=file_type,
    output_dir=output_dir, 
    project_name=project_name,
    scenarios=scenario_names,
    family_variables=family_variables,
    person_variables=person_variables,
    birth_year_range=birth_year_range,
    year_range=year_range
  )
```

